---
title: "Homework 5"
author: "YutingMei"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  github_document: default
  pdf_document: default
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

```{r}
library('MASS') ## for 'mcycle'
library('dplyr')
library('manipulate') ## for 'manipulate'
library(rlang)
library(caret)
```

* Randomly split the mcycle data into training (75%) and validation (25%) subsets.
```{r}
y <- mcycle$accel
x <- matrix(mcycle$times, length(mcycle$times), 1)
```

```{r}
plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
```

```{r}
dt = cbind(x,y)
```

```{r}
train_index = sample.int(nrow(dt), nrow(dt) * .75, replace = F)
index_list = seq(nrow(dt))
valid_index = index_list[!index_list %in% train_index]
train = dt[train_index, ]
valid = dt[valid_index, ]
```

* Using the mcycle data, consider predicting the mean acceleration as a function of time. Use the Nadaraya-Watson method with the k-NN kernel function to create a series of prediction models by varying the tuning parameter over a sequence of values. (hint: the script already implements this)

```{r}
## k-NN kernel function
## x  - n x p matrix of training inputs
## x0 - 1 x p input where to make prediction
## k  - number of nearest neighbors
kernel_k_nearest_neighbors <- function(x, x0, k) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
}
```

```{r}
## Make predictions using the NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## x0 - m x p matrix where to make predictions
## kern  - kernel function to use
## ... - arguments to pass to kernel function
nadaraya_watson <- function(y, x, x0, kern, ...) {
  k <- t(apply(x0, 1, function(x0_) {
    k_ <- kern(x, x0_, ...)
    k_/sum(k_)
  }))
  yhat <- drop(k %*% y)
  attr(yhat, 'k') <- k
  return(yhat)
}

```

```{r}
## Compute effective df using NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## kern  - kernel function to use
## ... - arguments to pass to kernel function
effective_df <- function(y, x, kern, ...) {
  y_hat <- nadaraya_watson(y, x, x,
    kern=kern, ...)
  sum(diag(attr(y_hat, 'k')))
}
```

```{r}
## loss function
## y    - train/test y
## yhat - predictions at train/test x
loss_squared_error <- function(y, yhat)
  (y - yhat)^2

## test/train error
## y    - train/test y
## yhat - predictions at train/test x
## loss - loss function
error <- function(y, yhat, loss=loss_squared_error)
  mean(loss(y, yhat))

## AIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
aic <- function(y, yhat, d)
  error(y, yhat) + 2/length(y)*d

## BIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
bic <- function(y, yhat, d)
  error(y, yhat) + log(length(y))/length(y)*d
```

```{r}
options(warn=-1)
## how does k affect shape of predictor and eff. df using k-nn kernel ?
# manipulate({
#   ## make predictions using NW method at training inputs
#   y_hat <- nadaraya_watson(y, x, x,
#     kern=kernel_k_nearest_neighbors, k=k_slider)
#   edf <- effective_df(y, x,
#     kern=kernel_k_nearest_neighbors, k=k_slider)
#   aic_ <- aic(y, y_hat, edf)
#   bic_ <- bic(y, y_hat, edf)
#   y_hat_plot <- nadaraya_watson(y, x, x_plot,
#     kern=kernel_k_nearest_neighbors, k=k_slider)
#   plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
#   legend('topright', legend = c(
#     paste0('eff. df = ', round(edf,1)),
#     paste0('aic = ', round(aic_, 1)),
#     paste0('bic = ', round(bic_, 1))),
#     bty='n')
#   lines(x_plot, y_hat_plot, col="#882255", lwd=2)
# }, k_slider=slider(1, 15, initial=3, step=1))
```

* With the squared-error loss function, compute and plot the training error, AIC, BIC, and validation error (using the validation data) as functions of the tuning parameter.

```{r}
error_combine = function(k_seq, y, x){
  y = matrix(y)
  x = matrix(x)
  aic_ = c()
  bic_ = c()
  error_ = c()
  for (i in k_seq){
  y_hat <- nadaraya_watson(y, x, x,
    kern=kernel_k_nearest_neighbors, k=i)
  edf <- effective_df(y, x,
    kern=kernel_k_nearest_neighbors, k=i)
  error_ = append(error_, error(y, y_hat))
  aic_ <- append(aic_, aic(y, y_hat, edf))
  bic_ <- append(bic_, bic(y, y_hat, edf))
  }
  data.frame(k = k_seq, aic = aic_, bic = bic_, error = error_)
}
```

```{r}
# train error
error_combine(seq(1,20), train[,2], train[,1])
```
```{r}
# test error
error_combine(seq(1,20), valid[,2], valid[,1])
```

* For each value of the tuning parameter, Perform 5-fold cross-validation using the combined training and validation data. This results in 5 estimates of test error per tuning parameter value.

```{r}
train.control <- trainControl(method = "cv", number = 5)
train_error_cv = train(y ~ .,
      method = 'knn',
      tuneGrid = expand.grid(k = 1:20),
      trControl = train.control,
      data = rbind(train, valid) %>% data.frame())
train_error_cv
```

* Plot the CV-estimated test error (average of the five estimates from each fold) as a function of the tuning parameter. Add vertical line segments to the figure (using the segments function in R) that represent one “standard error” of the CV-estimated test error (standard deviation of the five estimates from each fold).

```{r}
cv_all = train_error_cv$results %>%
  mutate(rmse_low = RMSE - 2*(RMSE / sqrt(train_error_cv$control$number)),
         rmse_high = RMSE - 2*(RMSE / sqrt(train_error_cv$control$number)))
```

```{r}
plot(x=range(as.matrix((cv_all$k))),
     y=range(range(cv_all$RMSE - cv_all$rmse_low, cv_all$RMSE + cv_all$rmse_high)),
     type='n',
     xlab='k',
     ylab='CV Test Error')
for(i in 1:nrow(cv_all)) {
  points(x=cv_all$k, y=cv_all$RMSE, pch=19, col='#00000055')
  lines(x=cv_all$k, y=cv_all$RMSE, col='#00000055')
  segments(x0 = cv_all$k, y0 = cv_all$RMSE - cv_all$rmse_low, x1 = cv_all$k, y1 = cv_all$rmse_high + cv_all$RMSE)
}

```

* Interpret the resulting figures and select a suitable value for the tuning parameter.
- When the k increases, the error become smaller for k from 1 to 15, but when k stil increases, the test error become greater again. k = 3 or 4 is suitable by one-standard-error rule.