---
title: "hw7"
author: "YutingMei"
date: "4/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library('rgl')
library('ElemStatLearn')
library('nnet')
library('dplyr')
library('keras')
```



```{r}
library(reticulate)

# create a new environment 
conda_create("r-reticulate")
```

```{r}
unlink("/Users/pritom/Library/r-miniconda", recursive = TRUE)
```

```{r}
remotes::install_github("rstudio/reticulate")
remotes::install_github("rstudio/tensorflow")
remotes::install_github("rstudio/keras")
reticulate::miniconda_uninstall()
reticulate::install_miniconda()
keras::install_keras()
```


```{r}
keras:::keras_version()
```


* Use the Keras library to re-implement the simple neural network discussed during lecture for the mixture data (see nnet.R). Use a single 10-node hidden layer; fully connected.

```{r}
data("mixture.example")
dat <- mixture.example
```

```{r}
plot_mixture_data <- function(dat=mixture.example, showtruth=FALSE) {
  ## create 3D graphic, rotate to view 2D x1/x2 projection
  par3d(FOV=1,userMatrix=diag(4))
  plot3d(dat$xnew[,1], dat$xnew[,2], dat$prob, type="n",
         xlab="x1", ylab="x2", zlab="",
         axes=FALSE, box=TRUE, aspect=1)
  ## plot points and bounding box
  x1r <- range(dat$px1)
  x2r <- range(dat$px2)
  pts <- plot3d(dat$x[,1], dat$x[,2], 1,
                type="p", radius=0.5, add=TRUE,
                col=ifelse(dat$y, "orange", "blue"))
  lns <- lines3d(x1r[c(1,2,2,1,1)], x2r[c(1,1,2,2,1)], 1)
  
  if(showtruth) {
    ## draw Bayes (True) classification boundary
    probm <- matrix(dat$prob, length(dat$px1), length(dat$px2))
    cls <- contourLines(dat$px1, dat$px2, probm, levels=0.5)
    pls <- lapply(cls, function(p) 
      lines3d(p$x, p$y, z=1, col='purple', lwd=3))
    ## plot marginal probability surface and decision plane
    sfc <- surface3d(dat$px1, dat$px2, dat$prob, alpha=1.0,
      color="gray", specular="gray")
    qds <- quads3d(x1r[c(1,2,2,1)], x2r[c(1,1,2,2)], 0.5, alpha=0.4,
      color="gray", lit=FALSE)
  }
}

## compute and plot predictions
plot_nnet_predictions <- function(fit, dat=mixture.example) {
  
  ## create figure
  plot_mixture_data()

  ## compute predictions from nnet
  preds <- predict(fit, dat$xnew, type="class")
  probs <- predict(fit, dat$xnew, type="raw")[,1]
  probm <- matrix(probs, length(dat$px1), length(dat$px2))
  cls <- contourLines(dat$px1, dat$px2, probm, levels=0.5)

  ## plot classification boundary
  pls <- lapply(cls, function(p) 
    lines3d(p$x, p$y, z=1, col='purple', lwd=2))
  
  ## plot probability surface and decision plane
  sfc <- surface3d(dat$px1, dat$px2, probs, alpha=1.0,
                   color="gray", specular="gray")
  qds <- quads3d(x1r[c(1,2,2,1)], x2r[c(1,1,2,2)], 0.5, alpha=0.4,
                 color="gray", lit=FALSE)
}

## plot data and 'true' probability surface
plot_mixture_data(showtruth=TRUE)
```


```{r}
## plot data and 'true' probability surface
plot_mixture_data(showtruth=TRUE)

## fit single hidden layer, fully connected NN 
## 10 hidden nodes
# fit <- nnet(x=dat$x, y=dat$y, size=10, entropy=TRUE, decay=0) 
model = keras_model_sequential()
model %>%
  layer_flatten() %>%
  layer_dense(units = 10, activation = 'relu') %>%
  layer_dense(units = 2, activation = 'softmax')
# plot_nnet_predictions(fit)
```

```{r}
model %>% compile(
  optimizer = 'adam', 
  loss = 'sparse_categorical_crossentropy',
  metrics = c('accuracy')
)
```

```{r}
model %>% fit(dat$x, dat$y, epochs = 5, verbose = 2)
```

```{r}
predictions <- model %>% predict(dat$xnew, type = "class")
```

```{r}
predictions %>% length
```

* Create a figure to illustrate that the predictions are (or are not) similar using the 'nnet' function versus the Keras model.

```{r}
fit <- nnet(x=dat$x, y=dat$y, size=10, entropy=TRUE, decay=0) 
```

```{r}
# result using nnet
predict_nn = predict(fit, dat$xnew, type = 'class')
```

```{r}
# result of keras
predictions_keras = apply(predictions, 1, function(x) ifelse(x[2] > x[1], 1, 0))
```

```{r}
plot_mix_data <- expression({
  plot(dat$xnew[,1], dat$xnew[,2],
       col=ifelse(predictions_keras==0, 'blue', 'orange'),
       pch=20,
       xlab=expression(x[1]),
       ylab=expression(x[2]),
       main = 'predictions from Keras')
})

eval(plot_mix_data)
```

```{r}
plot_mix_data <- expression({
  plot(dat$xnew[,1], dat$xnew[,2],
       col=ifelse(predict_nn=='0', 'blue', 'orange'),
       pch=20,
       xlab=expression(x[1]),
       ylab=expression(x[2]),
       main = 'predictions from nnet')
})

eval(plot_mix_data)
```

* (optional extra credit) Convert the neural network described in the "Image Classification" tutorial to a network that is similar to one of the convolutional networks described during lecture on 4/15 (i.e., Net-3, Net-4, or Net-5) and also described in the ESL book section 11.7. See the !ConvNet tutorial on the RStudio Keras website.

```{r}
# convert to net-3 with local connectivity
model <- keras_model_sequential() %>% 
  layer_conv_2d(filters = 32, kernel_size = c(3,3), activation = "relu", 
                input_shape = c(32,32,3)) %>% 
  layer_max_pooling_2d(pool_size = c(2,2)) %>% 
  layer_conv_2d(filters = 64, kernel_size = c(3,3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2,2)) %>% 
  layer_conv_2d(filters = 64, kernel_size = c(3,3), activation = "relu") %>%
  layer_locally_connected_2d(filters = 64, kernel_size = c(3,3), activation = "relu")
```

```{r}
summary(model)
```

```{r}
model %>% 
  layer_flatten() %>% 
  layer_dense(units = 64, activation = "relu") %>% 
  layer_dense(units = 10, activation = "softmax")
```

```{r}
summary(model)
```

