---
title: "Homework 2"
author: "YutingMei"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  github_document
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

```{r}
## load prostate data
prostate <- 
  read.table(url(
    'https://web.stanford.edu/~hastie/ElemStatLearn/datasets/prostate.data'))

## subset to training examples
prostate_train <- subset(prostate, train==TRUE)
```

```{r}
## plot lcavol vs lpsa
plot_psa_data <- function(dat=prostate_train) {
  plot(dat$lpsa, dat$lcavol,
       xlab="log Prostate Screening Antigen (psa)",
       ylab="log Cancer Volume (lcavol)",
       pch = 20)
}
plot_psa_data()

```

```{r}
############################
## regular linear regression
############################

## L2 loss function
L2_loss <- function(y, yhat)
  (y-yhat)^2

## fit simple linear model using numerical optimization
fit_lin <- function(y, x, loss=L2_loss, beta_init = c(-0.51, 0.75)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*x))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}

## make predictions from linear model
predict_lin <- function(x, beta)
  beta[1] + beta[2]*x

## fit linear model
lin_beta <- fit_lin(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=L2_loss)

## compute predictions for a grid of inputs
x_grid <- seq(min(prostate_train$lpsa),
              max(prostate_train$lpsa),
              length.out=100)
lin_pred <- predict_lin(x=x_grid, beta=lin_beta$par)

## plot data
plot_psa_data()

## plot predictions
lines(x=x_grid, y=lin_pred, col='darkgreen', lwd=2)

## do the same thing with 'lm'
lin_fit_lm <- lm(lcavol ~ lpsa, data=prostate_train)

## make predictins using 'lm' object
lin_pred_lm <- predict(lin_fit_lm, data.frame(lpsa=x_grid))

## plot predictions from 'lm'
lines(x=x_grid, y=lin_pred_lm, col='pink', lty=2, lwd=2)

```

* Write functions that implement the L1 loss and tilted absolute loss functions.

```{r}
L1_loss <- function(y, yhat)
  abs(y-yhat)

tilted_loss25 = function(y, yhat){
  l = y - yhat
  for (i in seq(1, length(l))){
    if (l[i] >= 0)
      l[i] = .25 * (l[i])
    else
      l[i] = (.25 - 1) * (l[i])
  }
  l
}

tilted_loss75 = function(y, yhat){
  l = y - yhat
  for (i in seq(1, length(l))){
    if (l[i] >= 0)
      l[i] = .75 * (l[i])
    else
      l[i] = (.75 - 1) * l[i]
  }
  l
}
```

* Create a figure that shows lpsa (x-axis) versus lcavol (y-axis). Add and label (using the 'legend' function) the linear model predictors associated with L2 loss, L1 loss, and tilted absolute value loss for tau = 0.25 and 0.75.

```{r}
fit_lin <- function(y, x, loss, beta_init = c(-0.51, 0.75)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*x))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}

process = function(loss){
  l_beta <- fit_lin(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=loss)

## compute predictions for a grid of inputs
x_grid <- seq(min(prostate_train$lpsa),
              max(prostate_train$lpsa),
              length.out=100)
predict_lin(x=x_grid, beta=l_beta$par)


}
```

```{r}
options(warn = -1)
pred_l1 = process(L1_loss)
pred_25 = process(tilted_loss25)
pred_75 = process(tilted_loss75)
```

```{r}
## plot data
plot_psa_data()

## plot predictions
lines(x=x_grid, y=pred_l1, col='orange', lwd=2)
lines(x=x_grid, y=lin_pred, col='green', lwd=2)
lines(x=x_grid, y=pred_25, col='red', lwd=2)
lines(x=x_grid, y=pred_75, col='blue', lwd=2)

legend("bottomright", c("L1 loss", "L2 loss", "tau = 0.25","tau = 0.75"), col = c("orange", "green", "red", "blue"),lty= c(1,1), lwd = 2)
```

* Write functions to fit and predict from a simple nonlinear model with three parameters defined by 'beta[1] + beta[2]*exp(-beta[3]*x)'. Hint: make copies of 'fit_lin' and 'predict_lin' and modify them to fit the nonlinear model. Use c(-1.0, 0.0, -0.3) as 'beta_init'.

```{r}
predict_nlin <- function(x, beta)
  beta[1] + beta[2]*exp(-beta[3]*x)

fit_nlin <- function(y, x, loss, beta_init = c(-1.0, 0.0, -0.3)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*exp(-beta[3]*x)))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}
```

```{r}
processn = function(loss){
  l_beta <- fit_nlin(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=loss)

## compute predictions for a grid of inputs
x_grid <- seq(min(prostate_train$lpsa),
              max(prostate_train$lpsa),
              length.out=100)
predict_nlin(x=x_grid, beta=l_beta$par)


}
```

```{r}
pred_l1n = processn(L1_loss)
pred_l2n = processn(L2_loss)
pred_25n = processn(tilted_loss25)
pred_75n = processn(tilted_loss75)
```

* Create a figure that shows lpsa (x-axis) versus lcavol (y-axis). Add and label (using the 'legend' function) the nonlinear model predictors associated with L2 loss, L1 loss, and tilted absolute value loss for tau = 0.25 and 0.75.

```{r}
## plot data
plot_psa_data()

## plot predictions
lines(x=x_grid, y=pred_l1n, col='orange', lwd=2)
lines(x=x_grid, y=pred_l2n, col='green', lwd=2)
lines(x=x_grid, y=pred_25n, col='red', lwd=2)
lines(x=x_grid, y=pred_75n, col='blue', lwd=2)

legend("bottomright", c("L1 loss", "L2 loss", "tau = 0.25","tau = 0.75"), col = c("orange", "green", "red", "blue"),lty= c(1,1), lwd = 2)
```

```{r}
##################################
## try modifying the loss function
##################################

## custom loss function
custom_loss <- function(y, yhat)
  (y-yhat)^2 + abs(y-yhat)

## plot custom loss function
err_grd <- seq(-1,1,length.out=200)
plot(err_grd, custom_loss(err_grd,0), type='l',
     xlab='y-yhat', ylab='custom loss')

## fit linear model with custom loss
lin_beta_custom <- fit_lin(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=custom_loss)

lin_pred_custom <- predict_lin(x=x_grid, beta=lin_beta_custom$par)

## plot data
plot_psa_data()

## plot predictions from L2 loss
lines(x=x_grid, y=lin_pred, col='darkgreen', lwd=2)

## plot predictions from custom loss
lines(x=x_grid, y=lin_pred_custom, col='pink', lwd=2, lty=2)




```

